Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-04-27 overleaf/paper_SIMD.tex
Did a pass on Section 4. Moving on to Seciton 5.

\begin{enumerate}
\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
in row-major order and at this point the programmer is responsible for linearization
and access. (This restriction can be easily lifted.)
\item Array subscrpts are plaintext values as specified by the rule:
\begin{semantics}
    \ntyperule{Array Access}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
}
\end{semantics}
The subscript $e$ is a function of the indices of the enclosing loops. 
For read access, the compiler allows an arbitrary such function. 
However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
%\item \ana{Add rules for logical ops}
%\item \ana{Add rules for arithmetic ops}
\item The final restriction involves MUX as expressed by the rule: 
\begin{semantics}
    \ntyperule{MUX}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
}
\end{semantics}
\end{enumerate}
Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
\begin{pythonn}
if e: A[i] = val
\end{pythonn}
Instead we had to write
\begin{pythonn}
if not(e): val = A[i];
A[i] = val;
\end{pythonn}

@@

The next steps construct the dependence graph and ???.

@@

\begin{itemize}
\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$. E.g., the def-use edge from \python{d = S[i,j] - C[j]} to \python{p = d*d} in the Biometric MPC-source is a same-level edge. A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge. E.g., $\texttt{min}_1 = \mathit{MUX}(\texttt{c},\texttt{sum}_1,\texttt{min}_1)$ to $\texttt{min}_0 = \phi(\texttt{min}_1, 10000)$ in Biometric is a same-level back-edge.
\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. 
%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$. E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. An inner-to-outer edge can be a back-edge as well in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).

@@

We require that all inputs are marked as either shared or plaintext.  We then determine if intermediate variables are shared through taint analysis with "taintedness" referring to the shared attribute.  Specifically, our compiler follows the following rules:
\begin{itemize}
\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is shared
\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext
\item Loop counters are always plaintext
\item Any variables which cannot be determined as shared or plaintext via the above rules are plaintext
\end{itemize}

The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  The final rule is needed to handle cycles of plaintext values.  For example, in the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and cannot be marked as plaintext through simple taint analysis:
{\small
\begin{verbatim}
plaintext_array = [0, 1, 2, ...]
sum!1 = 0
for i in range(0, N):
    sum!2 = PHI(sum!1, sum!3)
    sum!3 = sum!2 + plaintext_array[i]
\end{verbatim}
}
\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}

When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is implicitly converted to a shared value for that expression. \ben{Is this necessary to include?}

>>>>

\begin{enumerate}
\item Loops are of the form $0 <= i < I$ and bounds are fixed at compile time. 
It is a standard restriction in MPC that the bounds must be known at circuit-generation time.
\item Arrays are one-dimensional. N-dimensional arrays are linearized and accessed 
in row-major order and at this point the programmer is responsible for linearization
and access. (This restriction can be easily lifted.)
\item Array subscrpts are plaintext values as specified by the rule:
\begin{semantics}
    \ntyperule{Array Access}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e : \texttt{plain int} \quad \Gamma \vdash \texttt{A} : q \; \texttt{list}[\tau] \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{A[$e$]} : q \; \tau
}
\end{semantics}
The subscript $e$ is a function of the indices of the enclosing loops. 
For read access, the compiler allows an arbitrary such function. 
However, it restricts write access to \emph{canonical writes}, i.e., \texttt{A[i,j,k]} = ...
where \texttt{i}, \texttt{j} and \texttt{k} loop over the three dimensions of \texttt{A}.  
Write access such as for example \texttt{A[i,j+2]} = ... is not allowed. 
%\item \ana{Add rules for logical ops}
%\item \ana{Add rules for arithmetic ops}
\item The final restriction involves MUX as expressed by the rule: 
\begin{semantics}
    \ntyperule{MUX}{
%  \Gamma(\y) =  q_\y   \quad\quad \sigma(\f) = q_\f  \quad\quad \Gamma(\x) = q_\x \\
  \Gamma \vdash e_1 : q_1 \; \texttt{bool} \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \Gamma \vdash e_2 : q_2 \; \tau \quad \tau \in \{ \texttt{int}, \texttt{bool} \}
}{
  \Gamma \vdash \texttt{MUX($e_1$,$e_2$,$e_3$)} : q_1 \vee q_2 \vee q_3 \; \tau
}
\end{semantics}
\end{enumerate}
Importantly, the arguments of MUX are restricted to base types, as passing array arguments 
is not valid MOTION MPC. \ana{Check this.} This causes a slight inconvenience as we could not write 
\begin{pythonn}
if e: A[i] = val
\end{pythonn}
Instead we had to write
\begin{pythonn}
if not(e): val = A[i]
A[i] = val
\end{pythonn}

@@

The next steps construct the dependence graph and \ana{Add an intro here.}

@@

\begin{itemize}
\item same-level edge $X\rightarrow Y$ where $X$ and $Y$ are in the same loop nest, say $i,j,k$. E.g., the def-use edge 9 to 10 in the Biometric MPC Source in Listing~\ref{tab:source_and_MPC_source_and_optimized_MPC_source} is a same-level edge. A same-level edge can be a back-edge in which case a $\phi$ node is the target of the edge. 
E.g., 15 to 4 in Biometric is a same-level back-edge.
\item outer-to-inner $X\rightarrow Y$ where $X$ is in an outer loop nest, say $i$, and $Y$ is in an inner one, say $i,j,k$. E.g., 1 to 4 in Biometric forms is an outer-to-inner edge.
%A $\phi$-node can be a source or a target of an outer-to-inner forward edge.
\item inner-to-outer $X\rightarrow Y$ where $X$ is a $phi$-node in an inner loop nest, $i,j,k$, and $Y$ is in the enclosing loop nest $i,j$. E.g., the def-use from 8 to 12 gives rise to an inner-to-outer edge.
%E.g. $\texttt{sum}_0 = \phi(\texttt{sum}_1, 0)$ to $\texttt{c} = \mathit{CMP}(\texttt{sum}_0,\texttt{min}_0)$ is an inner-to-outer edge. 
An inner-to-outer edge can be a back-edge as well, in which case both $X$ and $Y$ are phi-nodes with the source $X$ in a loop nested into $Y$'s loop (not necessarily immediately).

@@

We require that all inputs are marked as either shared or plaintext, however, we do not require any other qualifier annotations in the program. Then we infer if intermediate variables and expressions are shared through taint analysis with "taintedness" referring to the shared attribute.  Specifically, the compiler follows the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one): 
\begin{enumerate}
\item Loop counters and array upper bounds are always \texttt{plain}.
\item If any variable on the right-hand side of an assignment is shared, then the assigned variable is \texttt{shared}.
%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
\end{enumerate}

%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
The final rule is needed in examples like the following. In the below snippet \texttt{sum!2} and \texttt{sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one. 
They are inferred as plaintext. 
%cannot be marked as plaintext through simple taint analysis:
{\small
\begin{verbatim}
plaintext_array = [0, 1, 2, ...]
sum!1 = 0
for i in range(0, N):
    sum!2 = PHI(sum!1, sum!3)
    sum!3 = sum!2 + plaintext_array[i]
\end{verbatim}
}
%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}

When converting to MOTION code, any plaintext value used in the right-hand side of a shared assignment is converted to a shared value for that expression. 
%\ben{Is this necessary to include?} \ana{Yes, that's important.}

Importantly, the taint analysis works on MPC Source, which lacks if-then-else control flow. 
This significantly simplifies the treatment as there is no need to handle conditionals and implicit flow.
